\documentclass{amsart}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{subfigure}

\theoremstyle{definition}
\newtheorem{mydef}{Definition}[section]
\newtheorem{lem}[mydef]{Lemma}
\newtheorem{thm}[mydef]{Theorem}
\newtheorem{cor}[mydef]{Corollary}
\newtheorem{claim}[mydef]{Claim}
\newtheorem{question}[mydef]{Question}
\newtheorem{hypothesis}[mydef]{Hypothesis}
\newtheorem{prop}[mydef]{Proposition}
\newtheorem{defin}[mydef]{Definition}
\newtheorem{example}[mydef]{Example}
\newtheorem{remark}[mydef]{Remark}
\newtheorem{notation}[mydef]{Notation}
\newtheorem{fact}[mydef]{Fact}

\newcommand{\Ls}{\mathcal{L}}

\newcommand{\Bb}{\mathcal{B}}
\newcommand{\Ps}{\mathcal{P}}

\newcommand{\Zz}{\mathbb{Z}}
\newcommand{\Qq}{\mathbb{Q}}
\newcommand{\Rr}{\mathbb{R}}
\newcommand{\Cc}{\mathbb{C}}


\newcommand{\Ff}{\mathbb{F}}
\newcommand{\Tor}{\operatorname{Tor}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\Ann}{\operatorname{Ann}}
\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\Mult}{\operatorname{Mult}}

\title[Math 123, Spring 2020: assignment 9]{Math 123 - Algebra II - Spring 2020 \\ Assignment 9}

%% Include only sections, not subsections, in the table of content.
\setcounter{tocdepth}{1}

\date{\today}

\begin{document}

%% No indentation at the start of each paragraph
%\parindent 0pt

\vspace*{-10em}
\maketitle

\textbf{Due Tuesday, April 7, 11h59pm.} (please submit your assignment as a PDF on Canvas). Unless otherwise noted, references are to Dummit and Foote, \emph{Algebra}, 3rd edition, Wiley, 2004.

\section*{Practice problems} (\emph{Not} for credit. No need to submit your solutions. Try to do them while minimally looking at the textbook)

\begin{enumerate}
\item Give the definition of the following: Frobenius map, perfect field, separable and inseparable degree, cyclotomic polynomial, automorphism, fixed field, Galois extension, Galois group. 
\item What is the degree of the cyclotomic field of $n$th root of unity? Prove that your answer is correct.
\item Prove that in a perfect field, irreducible polynomials are separable.
\item Prove that if $K / F$ is a finite extension, then $|\Aut (K / F)| \le [K : F]$. Give an example where equality holds, and an example where the inequality is strict.
\item Compute the Galois group of $\Qq (\sqrt{2}, \sqrt{3}) / \Qq$.
\item True or false?
  \begin{enumerate}
  \item If $f$ is a polynomial over a field of characteristic $p$ with zero derivative, then $f (x) = g (x^p)$ for some polynomial $g$.
  \item $\Qq (e^{2\pi i / n})$ has degree $n - 1$ over $\Qq$.
  \item $p$ is prime if and only if the $p$th cyclotomic polynomial is of the form $x^{p - 1} + x^{p - 2} + \ldots + x + 1$.
  \item If $K / F$ is Galois, and $L / K$ is Galois, then $L / F$ is Galois.
  \end{enumerate}
\end{enumerate}
\section*{Problems for credit}
\begin{enumerate}
\item You should have the assignment from another student available for review in your Canvas todo list. Review problem 3 from that assignment (submit your comment on Canvas, \emph{not} with this assignment). Refer to the peer review instructions on the course website for more details. \emph{You are encouraged not to look at the official solution before submitting your review!}

\item Let $F$ be a field and let $f (x) \in F[x]$ be a polynomial of degree $n \ge 1$. Assume that the characteristic of $F$ is either zero or at least $n$. Let $\alpha \in F$. Let $1 \le m \le n$. Show that $\alpha$ is a root of $f$ of multiplicity at least $m$ if and only if $f^{(k)} (\alpha)  = 0$ for all $k \in \{0, 1, 2, \ldots, m - 1\}$. 

  Here, $f^{(k)}$ denotes the $k$th derivative of $f$: $f^{(0)} = f$, and $f^{(k + 1)}$ is the derivative of $f^{(k)}$. 
\item
  \begin{enumerate}
  \item Show that any algebraically closed field is perfect.
  \item (DF, 13.6.4) Prove that if $n = p^k m$ where $p$ is prime and $m$ is coprime to $p$, then there are precisely $m$ distinct $n$th root of unity over a field of characteristic $p$.
  \item Show that a field $F$ is perfect if and only if any irreducible polynomial in $F[x]$ is separable.
  \end{enumerate}
\item
  \begin{enumerate}
  \item (DF, 14.1.2) Prove that the map $\tau: \Cc \to \Cc$ defined by $\tau (a + bi) = a- bi$ (called \emph{complex conjugation}) is an automorphism of $\Cc$.
  \item (DF, 14.1.3) Determine the fixed field of complex conjugation on $\Cc$.
  \item (DF, 14.1.4) Prove that $\Qq (\sqrt{2})$ and $\Qq (\sqrt{3})$ are not isomorphic.
  \end{enumerate}
\item Determine $\Aut (\Rr / \Qq)$. \emph{Hint: first show that any automorphism must preserve the ordering on $\Rr$ and be continuous.}

\end{enumerate}
\end{document}
